Recurrent event analysis in the presence of real-time high frequency data via random subsampling

### 2019-07-03 write-up/fda-recurrent.tex

\documentclass[11pt]{amsart}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows}
\usepackage{amsfonts,amssymb,amsthm, txfonts, pxfonts,amscd}
\usepackage[stretch=10]{microtype}
%\usepackage{hyperref}
\def\struckint{\mathop{%
\def\mathpalette##1##2{\mathchoice{##1\displaystyle##2}%
 {##1\textstyle##2}{##1\scriptstyle##2}{##1\scriptscriptstyle##2}}%
\mathpalette
{\vbox\bgroup\baselineskip0pt\lineskiplimit-1000pt\lineskip-1000pt
\halign\bgroup\hfill$}
{##$\hfill\cr{\intop}\cr\diagup\cr\egroup\egroup}%
}\limits}
\usepackage{natbib}
\usepackage{color}
\usepackage{booktabs,caption,fixltx2e}
\usepackage[flushleft]{threeparttable}

@@

\usepackage{soul,color, mathtools, fixmath}
\usepackage{multirow}
\def\pr{\mathop{\text{pr}}\nolimits}

@@

\title{Recurrent event analysis in the presence of real-time high
  frequency data via random subsampling}
\author{Walter Dempsey}
\address {Department of Statistics, 
  Harvard University, One Oxford Street
   Cambridge, MA  02138, USA}
 \email{wdempsey@fas.harvard.edu}

@@

The next step of our method is modeling~$\beta(s)$. Here, we leverage
ideas from the penalized spline literature~\citep{Ruppert2003,
  Wood2006book}. Let~$\mathbold{\phi} (s) = \{ \phi_1 (s), \ldots,
\phi_{K_b} (x) \}$ be a spline basis such that~$\beta(s) =
\sum_{j=1}^{K_b} b_j \phi_{j} (s) = \mathbold{\phi} (t) {\bf b}$
where~${\bf b} = [b_1, \ldots, b_{K_b}]^{\top}$. Thus, the integral
in~\eqref{eq:hazardlinear} can be restated as 
\begin{align*}
\int_{t-\Delta}^t X(s,t) \beta(s) ds 
  &= \int_{t-\Delta}^t \left[ \hat \mu(s,t) + {\bf c} (t)^\top
    \mathbold{\hat \psi} (s) \right] \times \left[
    \mathbold{\phi} (s) {\bf b} \right] \\
  &= [ M_{i,t}^\prime + C_{i,t}^\prime J_{\psi, \phi} ] {\bf b}
\end{align*}
where~$M_{i} (t) = (M_{i,1} (t), \ldots, M_{i,K_b} (t))$, $M_{i,j} (t)
= \int_{t-\Delta}^t \hat \mu (s,t) \phi_j (s)$, and~$J_{\psi, \phi}$ is a
$K_x \times K_b$ dimensional matrix with the~$(k,l)$th entry is equal
to~$\int_{0}^\Delta \psi_k (s) \phi_l (s)
ds$~\citep{RamsaySilverman2005}.

Given the basis for~$\beta(t)$, the model depends on choice of
both~$K_b$ and~$K_x$.  We follow~\cite{Ruppert2002} by choosing $K_b$
large enough to prevent undersmoothing and~$K_x \geq K_b$ to satisfy
identifability constraints.
In this paper, we follow the simple rule of thumb and set~$K_b = K_x =
35$.
As long as the choices of~$K_x$ and $K_b$ are large enough, their
impact on estimation is typically negligible. 
We formalize this statement in Theorem~\ref{thm:truncation}.
Note the choice of~$K_b$ large may lead to overfitting the data.
We therefore follow~\cite{GoldSmith2015} and induce smoothing by
assuming that~$\{ b_k \}_{k=3}^{K_b} \sim N(0, \sigma_b^2 I)$.
Below, we will exploit a connection between~\eqref{eq:approxscore}
and score equations for a logistic regression model.  Before moving
on, we introduce notation:
\begin{equation*}
h_i \left( t \given  H_{i,t}^{NX} ; \theta \right) \approx
\exp \left( Z_{i,t}^\top \gamma + g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
  + M_{i,t}^\top {\bf b} + C_{i,t}^\top J_{\phi, \psi} {\bf b} \right)
= \exp \left( W_{i,t}^\top \theta \right),
\end{equation*}
where~$\theta = (\gamma, \alpha, {\bf b})$ and~$\exp ( Z_{i,t}^\top
\gamma) =: h_0 (t)$ is the parametrized baseline intensity function.

@@

We next establish a connection between the above approximate score 
equations and the score equations for a logistic regression model.
All proofs are provided in the supplementary materials.
We can then exploit this connection to allow the model to be fit
robustly using standard mixed effects software~\citep{Ruppert2002,
  McCulloch2001}.

\begin{lemma} \normalfont
\label{lemma:logistic}
Under weights~\eqref{eq:waage_weights} and the log-linear intensity
function~\eqref{eq:hazardlinear}, the approximate score
function~\eqref{eq:approxscore} is equivalent to 
\[
\sum_{i=1}^n \sum_{t \in \bfT_i \cup D_i } \left[ {\bf 1}[t \in D_i]
  - \frac{1}{1 + \exp \left[- \left( \tilde{W}_{i,t}^\top \theta +
        \log \pi_i (t) \right) \right]} \right] \tilde{W}_{i,t} 
\]
where~$\tilde W_{i,t} = -W_{i,t}$. This is the score function for
logistic regression with binary response~$Y_i(t)$ for $t \in N_i \cup
D_i$ and~$i \in [n]$ where~$Y_i(t) = 1[t \in N_i]$, offset~$\log \pi_i
(t)$, and covariates~$\tilde W_{i,t}$. 
\end{lemma}

In this section, we make the following strong assumption.

\begin{assumption} \normalfont
\label{assumption:truncation}
Both the functional covariate~$\{ X(t,s) \}_{0\leq s \leq \Delta}$ and
the functional coefficient~$\beta(t)$ sit in the span of their
respective truncated basis expansions. That is, there
exists~$b^\star$ such that~$\beta(t) = \mathbold{\phi} (t) b^\star$
and~$X(s,t) = \hat \mu (s,t) + {\bf c}_i (t)^\top \mathbold{\hat \psi}
(s)$.
\end{assumption}

Assumption~\ref{assumption:truncation} allows us to ignore the impact
of truncation and present Lemma~\ref{lemma:simpleasym} which gives a
simple asymptotic theory in this particular case.

\begin{lemma} \normalfont
\label{lemma:simpleasym}
Under Assumption~\ref{assumption:truncation} and~$\Delta$ known, for
large~$n$ the estimate~$\hat \theta_n$ is approximately normal
\[
\sqrt{n} (\hat \theta - \theta) \to N(0, \Xi (\theta)^{-1})
\]
where
\[
  \Xi (\theta) = \int_{0}^{\tau} w(s; \theta) \left[ \frac{h^{(1)}(s;
      \theta) \times  h^{(1)} (s;\theta)^{\top}}{h(s; \theta)} \right]
  ds.
\]
and~$\tau$ is the random censoring time of the event process.
\end{lemma}
An estimator for~$\Xi(\theta)$ is
\begin{equation}
\label{eq:fisher}
  \hat \Xi (\theta) = n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup D_i} 
  w_i(t;\theta) (1-w_i(t;\theta)) \left[ \frac{h_i^{(1)}(t;
      \theta)}{h_i (t; \theta)} \right] \times  \left [
    \frac{h_i^{(1)} (t;\theta)}{h_i(t; \theta)} \right]^\top
\end{equation}
For the log-linear intensity model, the sampling-unbiased estimator
for~$\hat \Xi(\theta)$ is equivalent to the Fisher information for the
previously described logistic regression model.
This implies that subsampling from an inhomogeneous Poisson process,
standard logistic regression software can be used to fit the recurrent
event model by specifying an offset equal to~$\log \pi_i (t)$.
Not only this, Lemma~\ref{prop:optimal} shows
weights~\eqref{eq:waage_weights} are optimal within a particular class
of weighted estimating equations.

@@

\section{Theoretical analysis}

\begin{assumption}[Event process assumptions]
\begin{itemize}
\item The subsampling rate~$\pi_i (t) \in [L_\pi, U_\pi]$ 
\item There exists a nonnegative matrix
\item There exists 
\end{itemize}
\end{assumption}

\begin{assumption}[Covariance assumptions]
\begin{itemize}
\item 
\end{itemize}
\end{assumption}

\begin{assumption}[Functional data assumptions]
M1 -- M4 in 
\end{assumption}

\begin{thm}[No reconstruction error]
\label{thm:truncation}
Under regularity conditions A-E in Andersen et al. 1993, 
\[
\frac{n (\hat \theta - \theta) \Xi (\theta) (\hat \theta - \theta) -
  K_n+1}{\sqrt{2 (K_n + 1) }} \to N(0,1).
\]

\end{thm}

\begin{thm}[Reconstruction error OR $\hat \Delta > \Delta$]
\label{thm:biaseddown}

\end{thm}

@@

A core concern in mHealth is that sensor data can be missing due to
sensor wearing issue.
To see this, for the suicidal ideation case study, there are XX event
times across the whole dataset.  Of these, XX event times had complete 
data for the prior thirty minutes, XX had missingness less than 10\%
(i.e., sufficient to ), and XX had missingness above 30\%.

The issue here is that~$c_{i,k} (t)$ cannot be estimated if~$X(s,t)$
is not observed for all~$s \in [0,\Delta]$. 
Moreover, we want standard erros to reflect the uncertainty
in these coefficients when missing data is prevalent.
\cite{Goldsmith2015} suggest using best linear unbiased predictors
(BLUP) or posterior modes in the mixed effects model to
estimate~$c_{i,k} (t)$; however, one criticism would be when there is
substantial variability in these estimates.  To deal with this,
\cite{Crainiceanu2010, Goldsmith2010} take a full Bayesian analysis.
\cite{Yao2005} introduced PACE as a frequentist analysis that deals
with this issue.
However,~\cite{Reimherr2018} show that for sparse, irregular
longitudinal, the imputation model should not ignore the outcome~$Y_i
(t)$.

Here we present an extension of~\cite{Reimherr2018} to our setting,
leverage Lemma~\ref{lemma:logistic} and the marginal covariance
estimation procedure to build a multiple imputation approach.
\[
\E [ X_i (s,t) \given Y_i(t) = y, \bfx ] = \mu_y (s,t) + {\bf a}_{i,t}^\top
{\bf B}_i (\bfx_i (t) - \mathbold{\mu}_i )
\]
where~$\mu_y (t) = \E [ X_i (t) \given Y_i (t) = y]$, and
\[
{\bf a}_i^\top = ( \Sigma(s, )
\]
This generalizes the procedure of~\cite{Reimerr2018} to the
longitudinal data setting.
This is possible due to LEmma~\ref{} and our 
Here, we ignore the issue that~$X(t,s)$ is

@@

@@

\begin{proof}[Proof of Lemma~\ref{lemma:logistic}]
Recall~$\frac{d}{d\theta} \log h_i (t; \theta) = \frac{h^{(1)}_i (t;
  \theta)}{h_i (t; \theta)}$. Therefore, under the log-linear
intensity function, $\frac{d}{d\theta} \log h_i (t; \theta) =
W_{i,t}$. On the other hand,
\[
\frac{d}{d \theta} \log \left( \pi_i (t) + h_i (t;\theta) \right) = 
\frac{\exp \left( W_{i,t}^\prime \theta \right)}{\pi_i (t) + \exp
  \left( W_{i,t}^\prime \theta \right)} W_{i,t}
\]
Therefore,
\begin{align*}
\hat U_n (\theta) 
  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} W_{i,t} -
    \sum_{t \in N_i \cup D_i} \frac{\exp \left( W_{i,t}^\prime \theta
    \right)}{\pi_i (t) + \exp \left( W_{i,t}^\prime \theta \right)}
    W_{i,t} \right \} \\
  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} \frac{\pi_i (t)}{\pi_i (t)
    + \exp \left( W_{i,t}^\prime \theta \right)} W_{i,t} -
    \sum_{t \in D_i} \frac{\exp \left( W_{i,t}^\prime \theta
    \right)}{\pi_i (t) + \exp \left( W_{i,t}^\prime \theta \right)}
    W_{i,t} \right \} \\
  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} w_{i} (t; \theta) W_{i,t} -
    \sum_{t \in D_i} (1 - w_i (t; \theta)) W_{i,t} \right \} \\
  &= - \sum_{i=1}^n \sum_{t \in N_i \cup D_i} \left[ {\bf 1} [t \in
    D_i]  - w_{i} (t; \theta) \right] W_{i,t} \\
  &= \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
    \left[ {\bf 1} [t \in D_i]  - \frac{1}{1 + \exp\left( - (
          \tilde W_{i,t}^\top \theta + \log (\pi_i (t) ) ) \right)}
    \right] \tilde W_{i,t}.
\end{align*}
where~$\tilde W_{i,t} = - W_{i,t}$. This is exactly the score equation
for logistic regression with offset~$\log \pi_i (t)$.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lemma:simpleasym}]
Lemma~\ref{lemma:simpleasym} will be proven under regularity
conditions A-E in~\cite[pp. 420--421]{Andersen1993} along with the
following assumptions related to the subsampling rate~$\pi_i (t)$.
For simplicty we assume there exists~$\tau < \infty$ and all censoring
times satisfy~$\tau_i \leq \tau$. Define $R_i (t)$ to be the at risk
indicator for participant~$i$.  That is, 
\[
R_i (t) = \left \{ \begin{array}{c c} 1 & t \in (0,\tau_i) \\ 0
                                        & t > \tau_i\end{array} \right .
\]
We define the joint counting process
\[
M_i (t) = N_i (t) - \int_0^t (h_i (s; \theta) + \pi_i (s)) R_i (s) ds. 
\]
where~$N_i (t)$ is the counting process with jumps at all~$t \in
\bfT_i \cup D_i$.

>>>>

\documentclass[12pt]{amsart}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows}
\usepackage{amsfonts,amssymb,amsthm, txfonts, pxfonts,amscd}
\usepackage[stretch=10]{microtype}
%\usepackage{hyperref}
\def\struckint{\mathop{%
\def\mathpalette##1##2{\mathchoice{##1\displaystyle##2}%
 {##1\textstyle##2}{##1\scriptstyle##2}{##1\scriptscriptstyle##2}}%
\mathpalette
{\vbox\bgroup\baselineskip0pt\lineskiplimit-1000pt\lineskip-1000pt
\halign\bgroup\hfill$}
{##$\hfill\cr{\intop}\cr\diagup\cr\egroup\egroup}%
}\limits}
\usepackage{natbib}
\usepackage{color}
\usepackage{booktabs,caption,fixltx2e}
\usepackage[flushleft]{threeparttable}

@@

\usepackage{soul,color, mathtools, fixmath}
\usepackage{multirow}
\usepackage{enumitem}
\def\pr{\mathop{\text{pr}}\nolimits}

@@

\title[Recurrent event analysis with functional covariates via random
subsampling]{Recurrent event analysis in the presence of real-time
  high frequency data via random subsampling}
\author{Walter Dempsey}
\address {Department of Statistics, 
  Harvard University, One Oxford Street
   Cambridge, MA  02138, USA}
 \email{wdempsey@fas.harvard.edu}

@@

The next step of our method is modeling~$\beta(s)$. Here, we leverage
ideas from the penalized spline literature~\citep{Ruppert2003,
  Wood2006book}. Let~$\mathbold{\phi} (s) = \{ \phi_1 (s), \ldots,
\phi_{K_b} (x) \}$ be a spline basis and assume that~$\beta(s) =
\sum_{j=1}^{K_b} b_j \phi_{j} (s) = \mathbold{\phi} (t) {\bf b}$
where~${\bf b} = [b_1, \ldots, b_{K_b}]^{\top}$. Thus, the integral
in~\eqref{eq:hazardlinear} can be restated as 
\begin{align*}
\int_{t-\Delta}^t X(s,t) \beta(s) ds 
  &\approx \int_{t-\Delta}^t \left[ \hat \mu(s,t) + {\bf c} (t)^\top
    \mathbold{\hat \psi} (s) \right] \times \left[
    \mathbold{\phi} (s) {\bf b} \right] \\
  &= [ M_{i,t}^\prime + C_{i,t}^\prime J_{\hat \psi, \phi} ] {\bf b}
\end{align*}
where~$M_{i,t} = (M_{i,1,t}, \ldots, M_{i,K_b,t})$, $M_{i,j,t}
= \int_{t-\Delta}^t \hat \mu (s,t) \phi_j (s)$, and~$J_{\hat \psi,
  \phi}$ is a $K_x \times K_b$ dimensional matrix with the~$(k,l)$th
entry is equal to~$\int_{0}^\Delta \hat \psi_k (s) \phi_l (s)
ds$~\citep{RamsaySilverman2005}.

Given the basis for~$\beta(t)$, the model depends on choice of
both~$K_b$ and~$K_x$.  We follow~\cite{Ruppert2002} by choosing $K_b$
large enough to prevent undersmoothing and~$K_x \geq K_b$ to satisfy
identifability constraints.
While our theoretical analysis considers truncation levels that depend
on~$n$, in practice, we follow the simple rule of thumb and set~$K_b 
= K_x = 35$. As long as the choices of~$K_x$ and $K_b$ are large
enough, their impact on estimation is typically negligible. 
%We formalize this statement in Theorem~\ref{thm:truncation}.
Below, we will exploit a connection between~\eqref{eq:approxscore}
and score equations for a logistic regression model.  Before moving
on, we introduce notation:
\begin{equation}
\label{eq:approx_hazard}
h_i \left( t \given  H_{i,t}^{NX} ; \theta \right) \approx
\exp \left( Z_{i,t}^\top \gamma + g_t \left( H_{i,t}^{N} \right)^{\prime} \alpha
  + M_{i,t}^\top {\bf b} + C_{i,t}^\top J_{\hat \psi, \phi} {\bf b} \right)
= \exp \left( W_{i,t}^\top \theta \right),
\end{equation}
where~$\theta = (\gamma, \alpha, {\bf b})$ and~$\exp (
Z_{i,t}^\top \gamma) =: h_0 (t)$ is the parameterized baseline
intensity function. We write~$\tilde U_n (\theta)$ to denote
the approximate score function when substituting
in~\eqref{eq:approx_hazard} for~\eqref{eq:hazardlinear}.

@@

We next establish a connection between the above approximate score 
equations~$\tilde U_n (\theta)$ and the score equations for a logistic
regression model. We can then exploit this connection to allow the
model to be fit robustly using standard mixed effects
software~\citep{Ruppert2002, McCulloch2001}.

\begin{lemma} \normalfont
\label{lemma:logistic}
Under weights~\eqref{eq:waage_weights} and the log-linear intensity
function~\eqref{eq:approx_hazard}, the approximate score
function~$\tilde U_n (\theta)$ is equivalent to 
\[
\sum_{i=1}^n \sum_{t \in \bfT_i \cup D_i } \left[ {\bf 1}[t \in D_i]
  - \frac{1}{1 + \exp \left[- \left( \tilde{W}_{i,t}^\top \theta +
        \log \pi_i (t) \right) \right]} \right] \tilde{W}_{i,t} 
\]
where~$\tilde W_{i,t} = -W_{i,t}$. This is the score function for
logistic regression with binary response~$Y_i(t)$ for $t \in N_i \cup
D_i$ and~$i \in [n]$ where~$Y_i(t) = 1[t \in N_i]$, offset~$\log \pi_i
(t)$, and covariates~$\tilde W_{i,t}$. 
\end{lemma}

\subsection{Theoretical analysis}

Our theoretical analysis requires the several assumptions, some
regarding the subsampling procedure, some for the event process, and
some for the functional data. We state these assumptions and then our
main theorems.  
We start by assuming there exists a~$\tau < \infty$ such that all
individuals are no longer at risk (i.e.,~$\tau_i < \tau$ for all~$i$).
Moreover, define $R_i (t)$ to be the at risk indicator for
participant~$i$.  That is,   
\[
R_i (t) = \left \{ \begin{array}{c c} 1 & t \in (0,\tau_i) \\ 0
                                        & t > \tau_i\end{array} \right .
\]
Lemma~\ref{lemma:simpleasym} will be proven under regularity
conditions A-E in~\cite[pp. 420--421]{Andersen1993} along with the
following additional assumptions:

\begin{assumption}[Event process assumptions] 
\label{assumption:events}\normalfont
We assume the following holds:
\begin{enumerate}[label=(E.\arabic*)]
\item\label{E1} The subsampling rate is both lower and upper
  bounded for all ``at risk times''; that is,~$0 < L < \pi_i (t) < U <
  \infty$ for all~$i=1,2,\ldots$ and~$t \in [0,\tau]$. 
\item\label{E2} There exists a nonnegative definite matrix~$\Xi
  (\theta)$ such that
  \[
    n^{-1} \Xi_n (\theta) = n^{-1} \sum_{i=1}^n \int_0^\tau w_i (t;
    \theta) (1 - w_i (t; \theta)) \left[ \frac{h_i^{(1)}(t;
        \theta)}{h_i (t; \theta)} \right] \times  \left [
      \frac{h_i^{(1)} (t;\theta)}{h_i(t; \theta)} \right]^\top R_i (t)
    dt \overset{P}{\to} \Xi (\theta).
  \] 
\item\label{E3} There exists~$M$ such that $|W_{i,j,t}| < M$ for
  all~$(i,j,t)$.
\item\label{E4} For all~$j,k$ 
\[
n^{-1} \sum_{i=1}^n \int_0^\tau \left | \frac{d^2}{d\theta_j
    d\theta_k} h_i (t;\theta_0) \right|^2 R_i (t) dt \overset{P}{\to}
C < \infty
\]
as $n \to \infty$.
\end{enumerate}
\end{assumption}

We also require several assumptions due to the truncation of the
Karhunen-Lo{\`e}ve decomposition that represents~$X(t,s)$.
\begin{assumption}[Functional assumptions~\citep{Park2018}] \normalfont
\label{assumption:truncation}
The following assumptions are standard in prior work on longitudinal
functional data analysis~\citep{Park2018, Yao2005, ChenMuller2012}:
\begin{enumerate}[label=(A.\arabic*)]
\item\label{A1} $X = \{ X(s, T) : (s,T) \in \mathcal{S} \times
  \mathcal{T} \}$ is a square integrable element of the $L^2 (
  \mathcal{S} \times \mathcal{T})$
\item\label{A2} The sampling density~$g(T)$ is continuous
  and~$\sup |g(T)| < \infty$
\item\label{A3} $\E[X(s,T) X(s^\prime, T) X(s,T^\prime)
  X(s^\prime, T^\prime) ] < \infty$ for
  each $s,s^\prime [0,\Delta]$ and~$0 < T, T^\prime < \tau$.
\item\label{A4} $\E[\|X(\cdot, T)\|^4] < \infty$ for each~$0< T
  < \tau$.
\end{enumerate}
\end{assumption}
Assumption~\ref{A3} is guaranteed for by~\ref{E1} and~\ref{E3}, but is
stated here to emphasize its importance for the marginal covariance
procedure of~\cite{Park2018}.
Finally, for simplicity, we assume that there exists~$b^\star$
such that~$\beta(t) = \mathbold{\phi} (t) b^\star$; that is, the
true function~$\beta(t)$ sits in the span of the spline basis
expansion. 
With the necessary assumptions stated, we can now state
Lemma~\ref{lemma:simpleasym}, which provides asymptotic theory.
\begin{lemma} \normalfont
\label{lemma:simpleasym}
Under Assumption~\ref{assumption:events},
Assumption~\ref{assumption:truncation}, and~$\Delta$ known, for
large~$n$ the estimator~$\hat \theta_n$ is consistent; moreover,
\[
\sqrt{n} (\hat \theta - \theta) \to N(0, \Xi (\theta)^{-1})
\]
where
\[
  \Xi (\theta) = \int_{0}^{\tau} w(s; \theta) \left[ \frac{h^{(1)}(s;
      \theta) \times  h^{(1)} (s;\theta)^{\top}}{h(s; \theta)} \right]
  ds.
\]
and~$\tau$ is the random censoring time of the event process.
\end{lemma}
An estimator for~$\Xi(\theta)$ is
\begin{equation}
\label{eq:fisher}
  \hat \Xi (\theta) = n^{-1} \sum_{i=1}^n \sum_{t \in \bfT_i \cup D_i} 
  w_i(t;\theta) (1-w_i(t;\theta)) \left[ \frac{h_i^{(1)}(t;
      \theta)}{h_i (t; \theta)} \right] \times  \left [
    \frac{h_i^{(1)} (t;\theta)}{h_i(t; \theta)} \right]^\top
\end{equation}
For the log-linear intensity model, the sampling-unbiased estimator
for~$\hat \Xi(\theta)$ is equivalent to the Fisher information for the
previously described logistic regression model.
This implies that subsampling from an inhomogeneous Poisson process,
standard logistic regression software can be used to fit the recurrent
event model by specifying an offset equal to~$\log \pi_i (t)$.
Not only this, Lemma~\ref{prop:optimal} shows
weights~\eqref{eq:waage_weights} are optimal within a particular class
of weighted estimating equations.

@@

\subsection{Penalized functional regression models}

Recall theoretical results were proven under the assumption that there
exists~$b^\star$ such that~$\beta(t) = \phi(t) {\bf b}^\star$.
To make this assumption plausible, we set~$K_b$ large enough (but less
than~$K_x$ to ensure identifiability) such that the spline basis
expansion is sufficiently expressive. However, in practice, such a
choice of~$K_b$ may lead to overfitting the data. We therefore, in the
remainder of this paper, follow~\cite{GoldSmith2015} and induce
smoothing by assuming that~$\{ b_k \}_{k=3}^{K_b} \sim N(0, \sigma_b^2
I)$. Incorporating penalization is simple due to the connection
between with the score functions for logistic regression.

\subsection{Confidence intervals for~$\beta(t)$}

Due to the connection with generalized mixed effects models, we can
leverage existing inferential machinery to obtain variance-covariance
estimates of model parameters. 
That is, if~$\hat \Sigma_{bb}$ is the $K_b \times K_b$ deimnsional
matrix obtained by plugging in the REML estimates of variance
components into the formula for the variance of~$\hat b$, then the
standard error for estimate at time~$t_0$ -- i.e., ~$\hat \beta (t_0)
= \phi(t_0) \hat {\bf b}$ -- is given by~$\sqrt{ \phi (t_0 )
  \hat \Sigma_{bb} \phi(t_0)^\top}$.  Then the approximate 95\%
confidence interval can be constructed as~$\hat \beta (t_0) \pm 1.96
\sqrt{\phi (t_0) \hat \Sigma_{bb} \phi(t_0)^\top}$.

We acknowledge two important limitations of confidence intervals
obtained via this approach. First, penalization may lead to confidence
intervals that perform poorly in regions where~$\hat \beta(t)$ is
oversmoothed. Second, we ignore the variability inherent in the
longitudinal functional principal component analysis; that is, our
estimates ignore the variability in estimation of eigenfunctions~$\hat
\psi$ as well as the coefficients~$\hat c_{i,k}(t)$.
Joint modeling could be considered as in~\cite{Crainiceanu2010},
however, this is beyond the scope of this article.

@@

A core concern in mHealth is that sensor data can be missing due to
sensor wearing issue.
To see this, for the suicidal ideation case study, there are 2139
self-identified moments of distress across all 91 participants.  Of
these, 1289 event times had complete data for the prior thirty minutes,
1984 had fraction of missing data on a fine grid less than 30\%, and
1998 had fraction of missing data on a fine grid less than 10\%.

Missing data is a critical issue because~$c_{i,k} (t)$ cannot be
estimated if~$X(s,t)$ is not observed for all~$s \in [0,\Delta]$. 
Moreover, we want standard errors to reflect the uncertainty
in these coefficients when missing data is prevalent.
\cite{Goldsmith2015} suggest using best linear unbiased predictors
(BLUP) or posterior modes in the mixed effects model to
estimate~$c_{i,k} (t)$; however, one criticism would be when there is
substantial variability in these estimates.  To deal with this,
\cite{Crainiceanu2010, Goldsmith2010} take a full Bayesian analysis.
\cite{Yao2005} introduced PACE as a frequentist analysis that deals
with this issue.
However,~\cite{Reimherr2018} show that for sparse, irregular
longitudinal, the imputation model should not ignore the outcome~$Y_i
(t)$.

Here we present an extension of~\cite{Reimherr2018} to our setting,
leverage Lemma~\ref{lemma:logistic} and the marginal covariance
estimation procedure to build a multiple imputation approach.
\[
\E [ X_i (s,t) \given Y_i(t) = y, \bfx ] = \mu_y (s,t) + {\bf a}_{i,t}^\top
{\bf B}_i (\bfx_i (t) - \mathbold{\mu}_i )
\]
where~$\mu_y (t) = \E [ X_i (t) \given Y_i (t) = y]$, and
\[
{\bf a}_i^\top = ( \Sigma(s, )
\]
This generalizes the procedure of~\cite{Reimerr2018} to the
longitudinal data setting.
This is possible due to  and our 
Here, we ignore the issue that~$X(t,s)$ is

@@

\subsection{Coverage properties}

@@

\begin{proof}[Proof of Lemma~\ref{lemma:logistic}]
Recall~$\frac{d}{d\theta} \log h_i (t; \theta) = \frac{h^{(1)}_i (t;
  \theta)}{h_i (t; \theta)}$. Under the log-linear
intensity function given by~\eqref{eq:approx_hazard},
$\frac{d}{d\theta} \log h_i (t; \theta) = W_{i,t}$ and 
\[
\frac{d}{d \theta} \log \left( \pi_i (t) + h_i (t;\theta) \right) = 
\frac{\exp \left( W_{i,t}^\top \theta \right)}{\pi_i (t) + \exp
  \left( W_{i,t}^\top \theta \right)} W_{i,t}
\]
Therefore,
\begin{align*}
\tilde U_n (\theta) 
  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} W_{i,t} -
    \sum_{t \in N_i \cup D_i} \frac{\exp \left( W_{i,t}^\top \theta
    \right)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)}
    W_{i,t} \right \} \\
  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} \frac{\pi_i (t)}{\pi_i (t)
    + \exp \left( W_{i,t}^\top \theta \right)} W_{i,t} -
    \sum_{t \in D_i} \frac{\exp \left( W_{i,t}^\top \theta
    \right)}{\pi_i (t) + \exp \left( W_{i,t}^\top \theta \right)}
    W_{i,t} \right \} \\
  &= \sum_{i=1}^n \left \{ \sum_{t \in N_i} w_{i} (t; \theta) W_{i,t} -
    \sum_{t \in D_i} (1 - w_i (t; \theta)) W_{i,t} \right \} \\
  &= - \sum_{i=1}^n \sum_{t \in N_i \cup D_i} \left[ {\bf 1} [t \in
    D_i]  - w_{i} (t; \theta) \right] W_{i,t} \\
  &= \sum_{i=1}^n \sum_{t \in N_i \cup D_i}
    \left[ {\bf 1} [t \in D_i]  - \frac{1}{1 + \exp\left( - (
          \tilde W_{i,t}^\top \theta + \log (\pi_i (t) ) ) \right)}
    \right] \tilde W_{i,t}.
\end{align*}
where~$\tilde W_{i,t} = - W_{i,t}$. This is exactly the score equation
for logistic regression with offset~$\log \pi_i (t)$.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lemma:simpleasym}]
Lemma~\ref{lemma:simpleasym} will be proven under regularity
conditions A-E in~\cite[pp. 420--421]{Andersen1993} along with the
following assumptions related to the subsampling rate~$\pi_i (t)$.
We define the joint counting process
\[
M_i (t) = N_i (t) - \int_0^t (h_i (s; \theta) + \pi_i (s)) R_i (s) ds. 
\]
where~$N_i (t)$ is the counting process with jumps at all~$t \in
\bfT_i \cup D_i$.

Asymptotic consistency is guaranteed~\cite[Theorem VI.1.1]{Andersen1993} if
\begin{equation}
\label{eq:andersen1}
n^{-1} \tilde U_n (\theta_0 ) \overset{P}{\to} 0,
\end{equation}
and 
\begin{equation}
\label{eq:andersen2}
n^{-1} \frac{d}{d \theta^{\top}} \tilde U_n (\theta) \big |_{\theta =
  \theta_0} = \Xi (\theta),
\end{equation}
and
\begin{equation}
\label{eq:andersen3}
\lim_{n \to \infty} P \left( \left | n^{-1} \frac{d^2}{d \theta_j d
      \theta_k} \tilde U_n (\theta) \right | < M \text{ for all }
  j,k \text{ and all } \theta \in \Theta_0 \right) = 1
\end{equation}
To prove~\eqref{eq:andersen1}, we decompose~$\hat U_n (\theta_0)$ into
three terms
\[
n^{-1} \tilde U_n (\theta_0) = n^{-1} U_n (\theta_0 )  + n^{-1} \left(
  \hat U_n (\theta_0) - U_n (\theta) \right) + n^{-1} \left(\tilde U_n
  (\theta_0) - \hat U_n (\theta_0) \right)
\]
where, recall,~$\hat U_n (\theta_0)$ are the logistic score equations
with $\int_0^\infty X_i(t-s) \beta(s) ds$ and~$\tilde U_n (\theta_0)$
with the approximation~$C_{i,t}^\top J_{\phi, \psi} {\bf b}$. 
\cite{Andersen1993} (Theorem VI.1.1) show that~$n^{-1} U_n (\theta_0)
\to 0$. The second term
\begin{align*}
n^{-1} \left( \hat U_n (\theta_0) - U_n (\theta_0) \right) =
  \frac{1}{n} \sum_{i=1}^n \int_0^{\tau} \frac{h^{(1)} (s;
  \theta_0)}{\pi_i (s) + h (s;\theta_0)} d M_i (s). 
\end{align*}
where~$h(\cdot; \theta_0)$ is given by~\eqref{eq:hazardlinear}.
Lenglart's inequality \walt{check} implies the second term converges
in probability to zero under Assumptions~\ref{A1} and~\ref{A2}.
